The expected shape of the Milky Way's Dark Matter halo

### 2018-09-18 Paper/paper.tex
Bib complete for now

Although observations predict more symmetric halos than DM-only simulations, there is strong evidence claiming that the presence of baryons need/make more axisymmetrical halos. Some studies have shown that to ensure the stability of a hydrodynamical (numeric) disk artificially embeded in a static(non-changing) DM halo, its shape must be axisymmetrical \cite{Scannapieco_et_al._2009,Yurin_and_Springel_2015}. Other works have studied this rounding effect by emulating a disk as an adiabatically growing potential inside an N-body triaxial DM halo \cite{Debattista_et_al._2008,Debattista-Roskar, katzantzidis}, finding that (findings of this study -debattista-). Other works of this type?\\

While these studies are elucidating about dynamical aspects that may play important roles in the rounding of the DM halo shape, they do not follow baryons in the whole cosmological context. Abadi-Navarro and Bryan-Kay studied the evolution of galaxies in hydrodynamical simulations with restricted models of feedback \cite{Abadi-Navarro,Bryan-Kay} finding that the galaxy formation efficiency \cite{Abadi} and the history of formation may also determine the extent of this rounding.\\

@@

There is no trivial way to calculate the DM halo shape at a determined radius due to the discretization of particles. For practical effects, observational models focus on the shape of the DM isopotential \cite{isopotential} or isodensity \cite{isodensity} contours. In cosmological simulations, it is preferred to work with the isodensity contours which is directly calculated from particle positions. However, these density contours are not smooth and are very sensitive to the presence of small satelites \cite{Springel Isodensity}. For this reason we determine the shape taking into account volume-enclosed particles, rather than shell-enclosed. While this approach allows the cumulative influence of outter shells on the outter shapes, it has been demonstrated for DM-only simulations that this effect becomes important only for big radii \cite{vera-Ciro aquarius}.  We therefore follow the method of \cite[Allgood et al. 2006]{Allgood_et_al._2006} that consits in the use of the reduced inertia tensor,   \\

@@

\subsection{The radial tendency of axial ratios}
To have a more complete picture about the effect of baryons on the DM halo shape, we begin by verifying/understanding the radial evolution (dependence?) of the shape. In the case of DM-only simulations, it is expected that halos get monotonically rounder at larger radii due to (the gravitational effect of inner shells?) \citep{Vera-Ciro et al. 2012}. A graphic example of this rounding can be checked on figure \ref{fig:slices}, where we plot the a projection of a DM-only halo aligned with the calculated minor axis at two different radius together with the estimated projected ellipsoidal. Formally, this effect is reflected in the axial ratios $b/a,c/a$, which tend to unity on the outerskirts of the halo. As a representative sample of this relations, we show on figure \ref{fig:DM relation} the evolution, for a specific halo, of the axial ratios $b/a,c/a$ and the triaxiality parameter $T=\frac{a^2-b^2}{a^2-c^2}$ in a wide range of radii.\\

@@

As a check of consistency, we compare our new isopotential shape results with the analytic expression of the $(1-q_{\phi})\frac{1}{3}(1-q_{\rho})$ \citep[Binney and Tremaine year]{Binney and tremaine}, taking the volume-enclosed axial ratios as an approximation for the isodensity contour ratios $q_{\rho}$. Although this analytic expression if meant to be used for logarithmic axisymmetric halos, it works well as a first approximation for nearly axisymmetric halos as those produced by our disced galaxies. We find that the difference between the real and the analytic isopotential axial ratios is not bigger than $quantity percent$. With this, we may now present on table \ref{tabe:isopotential} our observationally-comparable results for MHD halos at two important radii one corresponding to the approximate regimes where the MW DM halo is usually constrained.\\

>>>>

Although observations predict more symmetric halos than DM-only simulations, there is strong evidence claiming that the presence of baryons need/make more axisymmetrical halos. Some studies have shown that to ensure the stability of a hydrodynamical (numeric) disk artificially embeded in a static(non-changing) DM halo, its shape must be axisymmetrical \cite{Scannapieco_et_al._2009,Yurin_and_Springel_2015}. Other works have studied this rounding effect by emulating a disk as an adiabatically growing potential inside an N-body triaxial DM halo \cite{Debattista_et_al._2008,Kazantzidis_et_al._2010,Debattista_et_al._2013}, finding that (findings of this study -debattista-). Other works of this type?\\

While these studies are elucidating about dynamical aspects that may play important roles in the rounding of the DM halo shape, they do not follow baryons in the whole cosmological context. Abadi et al. and Bryan et al. studied the evolution of galaxies in hydrodynamical simulations with restricted models of feedback \cite{Abadi_et_al._2010,Bryan_et_al._2013} finding that the galaxy formation efficiency \cite{Abadi} and the history of formation may also determine the extent of this rounding.\\

@@

There is no trivial way to calculate the DM halo shape at a determined radius due to the discretization of particles. For practical effects, observational models focus on the shape of the DM isopotential \cite{Law_and_Majewski_2010} or isodensity \cite{Loebman_et_al._2012} contours. In cosmological simulations, it is preferred to work with the isodensity contours which is directly calculated from particle positions. However, these density contours are not smooth and are very sensitive to the presence of small satelites \cite{Springel Isodensity}. For this reason we determine the shape taking into account volume-enclosed particles, rather than shell-enclosed. While this approach allows the cumulative influence of outter shells on the outter shapes, it has been demonstrated for DM-only simulations that this effect becomes important only for big radii \cite{Vera-Ciro_et_al._2011}.  We therefore follow the method of \cite[Allgood et al. 2006]{Allgood_et_al._2006} that consits in the use of the reduced inertia tensor,   \\

@@

\subsection{The radial tendency of axial ratios}
To have a more complete picture about the effect of baryons on the DM halo shape, we begin by verifying/understanding the radial evolution (dependence?) of the shape. In the case of DM-only simulations, it is expected that halos get monotonically rounder at larger radii due to (the gravitational effect of inner shells?) \citep{Vera-Ciro_et_al._2012}. A graphic example of this rounding can be checked on figure \ref{fig:slices}, where we plot the a projection of a DM-only halo aligned with the calculated minor axis at two different radius together with the estimated projected ellipsoidal. Formally, this effect is reflected in the axial ratios $b/a,c/a$, which tend to unity on the outerskirts of the halo. As a representative sample of this relations, we show on figure \ref{fig:DM relation} the evolution, for a specific halo, of the axial ratios $b/a,c/a$ and the triaxiality parameter $T=\frac{a^2-b^2}{a^2-c^2}$ in a wide range of radii.\\

@@

As a check of consistency, we compare our new isopotential shape results with the analytic expression of the $(1-q_{\phi})\frac{1}{3}(1-q_{\rho})$ \citep[Binney and Tremaine 2008]{Binney_and_Tremaine_2008}, taking the volume-enclosed axial ratios as an approximation for the isodensity contour ratios $q_{\rho}$. Although this analytic expression if meant to be used for logarithmic axisymmetric halos, it works well as a first approximation for nearly axisymmetric halos as those produced by our disced galaxies. We find that the difference between the real and the analytic isopotential axial ratios is not bigger than $quantity percent$. With this, we may now present on table \ref{tabe:isopotential} our observationally-comparable results for MHD halos at two important radii one corresponding to the approximate regimes where the MW DM halo is usually constrained.\\
